---
title: "corona"
author: "ipi"
date: "3/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(WDI)
library(readxl)
library(httr)
library(knitr)
library(tidyverse)
library(estimatr)
```

# Options

```{r}
download_corona_data <- FALSE
download_wdi_data    <- FALSE
```

# Get Corona Data

```{r, eval = download_corona_data}
url <- paste("https://www.ecdc.europa.eu/sites/default/files/documents/COVID-19-geographic-disbtribution-worldwide-",format(Sys.time(), "%Y-%m-%d"), ".xlsx", sep = "")
GET(url, authenticate(":", ":", type="ntlm"), write_disk(tf <- tempfile(fileext = ".xlsx")))
df <- read_excel(tf)

write.csv(df, "saved/corona.csv")
```

```{r}
corona <- read.csv("saved/corona.csv")
head(corona) %>% kable
```

# Get WDI data

```{r, eval = download_wdi_data}

# write.csv(WDI_data$series,  "saved/wdi_series.csv")

wdi <- WDI(country = "all", 
         indicator = c(
            goveffect = "GE.EST", 
            ineq = "SI.DST.10TH.10", 
            gdppc = "NY.GDP.PCAP.KD",
            poptot = "SP.POP.TOTL",
            popgrowth = "SP.POP.GROW",
            older_m =  "SP.POP.65UP.MA.IN",
            older_f = "SP.POP.65UP.FE.IN",
            gini = "SI.POV.GINI"
          ), 
  start = 2018, end = 2018, extra = FALSE, cache = NULL)

write.csv(wdi, "saved/wdi.csv")
```

```{r}

wdi <- read.csv("saved/wdi.csv")

head(wdi) %>% kable()


```

# Merge

```{r}
df <- left_join(corona, wdi, by = c("GeoId" = "iso2c")) %>%
      mutate(share_older = (older_m + older_f)/poptot, 
             date = as.Date(DateRep),
             elapsed = -2 + date - min(date))

```


# Simple plots

```{r}

with(df %>% filter(elapsed == 80 & GeoId != "CN"), {
  par(mfrow = c(2,2))
  
  plot(share_older, Cases, type = "n", xlab = "Share of population over 65")
  text(share_older, Cases, GeoId)

  plot(gdppc, Cases, type = "n", xlab = "Per capita GDP")
  text(gdppc, Cases, GeoId)

  plot(goveffect, Cases, type = "n", xlab = "Government effectiveness")
  text(goveffect, Cases, GeoId)
  
  })

```


# Analysis: Patterns changing over time
Note: models exclude China an focuses on spread countries only.

Expecting a shift to poor places over time -- though of course burden still growingin richest places. 

Prediction is that this will fall in the future as OECD gains control and LDCs lose control.

```{r, warning = FALSE, message = FALSE}

results <- function(Y = "Cases", Xs = "gdppc + poptot + share_older + goveffect", range = 50:83, data = filter(df, GeoId != "CN")){
  my_formula <- paste0(Y, " ~", Xs)
  lapply(range, function(j) {
    df <- filter(data, elapsed == j)
    lm_robust(as.formula(my_formula), data = df) %>% 
      tidy %>% select(term, estimate, conf.low, conf.high) %>% 
      mutate(elapsed = j) }) %>%
  bind_rows() %>%
  data.frame() %>%
  filter(term != "(Intercept)")
}


```


# Plot

```{r, warning = FALSE}

do_plots <- function(df)
  df %>% 
    ggplot(aes(x=elapsed, y=estimate)) + 
      geom_point() +
      geom_line() +
      geom_errorbar(aes(ymin=conf.low, ymax=conf.high), colour="black", width=.1) +
      xlab("Days since 1 Jan 2020") +
      ylab("Coefficient") +
      theme_bw()  +
      facet_wrap(~term, scales = "free")
```

## Cases

```{r, warning = FALSE, message = FALSE}
results_cases <- results("Cases")
do_plots(results_cases)
```

## Deaths

```{r, warning = FALSE, message = FALSE}
results_deaths <- results("Deaths")
do_plots(results_deaths)
```

## To do

* Bring in political variables
  * Democracy
  * L/R orientation?
  * Media?
  * Ethnic Fragmentation: Prediction is it slows down onset but also slows responses
  
* Outcomes:
  * Cases
  * Deaths
  * Policies
  * Growth rate of cases
  * Growth rate of deaths